---
title: "Worked Example Using the IST Dataset"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Worked Example Using IST Dataset}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction
This R Markdown document illustrates example usage of the RESIDE package using the IST dataset.

# Setup
Load the RESIDE package and set a seed for reproducibility and store the folder directory for export / import.
```{r load-library}
# Load the Library
library(RESIDE)
# Set the seed
set.seed(1234)
# Store the folder path used for import / export
folder_path <- tempdir()
```

# Summarise original data
Select the variables of interest and summarise. Selection of variables is optional, if the variables are known. They may not be known until the marginal distributions have been received.
```{r summarise-data}
# Select variables of interest from the IST dataset.
IST_original <- IST |> dplyr::select(
  AGE, # AGE at Randomisation
  SEX, # SEX M/F
  RATRIAL, # Atrial Fibrillation Y/N at Randomisation 
  # (not coded for 984 patients in the pilot phase)
  RSBP, # Systolic Blood Pressure at Randomisation
  STRK14 # Indicator of Any Stroke at 14 days
)

# Convert the character variables to factors (to allow for summary)
IST_original <- IST_original |> dplyr::mutate_if(is.character, factor)

# Produce a summary of the variables
summary(IST_original)

```

# Fit a cox model on the original data
```{r original-model}
# Load survival and dplyr libraries
library(survival) # For Cox PH model
library(dplyr) # For data manipulation

# Stroke event is measured at 14 days, so set this for patients
IST_original$DAY <- 14

# Illustrate the 984 missing values
sum(IST_original$RATRIAL == "")

# Remove the missing values
IST_original <- IST_original[!IST_original$RATRIAL == "",]

# Drop the factor name for the missing values
IST_original$RATRIAL <- droplevels(IST_original$RATRIAL)

# Summarise the variable to show there are no longer missing values
summary(IST_original$RATRIAL)

# Fit a Cox PH model
cox.ph <- coxph(Surv(DAY, STRK14) ~ AGE + SEX + RATRIAL + RSBP, data = IST_original) 

# Output the summary of the Cox PH Model
cox.ph
```

# Get the marginal distributions
Use the `get_marginal_distributions()` function to get the marginal distributions, additionally selecting which variables using the `variables` parameter.
```{r get-marginals}
# Get the Marginal Distributions for the selected variables
marginals <- get_marginal_distributions(
  IST,
  variables = c(
    "AGE",
    "SEX",
    "RATRIAL",
    "RSBP",
    "STRK14"
  )
)

```

# Export the marginal distributions
Export the marginal distributions using the `export_marginal_distributions()` function, using the `force` parameter to override any existing files.
```{r export-marginals}
# Export the Marginal Distributions
export_marginal_distributions(marginals,
                              folder_path = folder_path,
                              force = TRUE)
```
# Reimport marginal distributions
Import the exported marginal distributions using the `import_marginal_distributions()` function
```{r import-marginals}
# Import the Marginal Distributions
imported_marginals <- import_marginal_distributions(folder_path = folder_path)
```

# Synthesise data from marginal distributions (without correlations)
Synthesise data from the imported marginals using the `synthesise_data` function.
```{r synthesis-data}
# Synthesise a dataset from the imported Marginal Distributions (without correlations)
sim_df <- synthesise_data(imported_marginals)
```

# Summarise the synthesised data
Summarise the simulated data
```{r summarise-sim-data}
# Convert any Character variables to Factors
sim_df <- sim_df |> dplyr::mutate_if(is.character, factor)
# Summarise the synthesised data
summary(sim_df)
```

# Fit a cox model on the simulated data
Fit the same cox model as earlier except this time on the simulated data.
```{r sim-data-model}

# As before the events are measured at day 14
sim_df$DAY <- 14

# Show that the missing observations are in the data
sum(sim_df$RATRIAL == "")

# Remove the missing observations
sim_df <- sim_df[!sim_df$RATRIAL == "",]

# Remove the missing factor name
sim_df$RATRIAL <- droplevels(sim_df$RATRIAL)

# Show that there are no missing observations
summary(sim_df$RATRIAL)

# Fit the model on the synthesised data
cox.ph.sim <- coxph(Surv(DAY, STRK14) ~ AGE + SEX + RATRIAL + RSBP, data = sim_df) 

# Show a summary of the model
cox.ph.sim
```

# Synthesise data with correlations
Synthesise data from the imported marginals with correlations, by exporting an empty correlation matrix, then reimporting it and using the `correlation_matrix` parameter to specify the correlation matrix. Note the matrix was edited in R for this example but users would be expected to edit the correlation matrix in a program such as excel.
```{r synthesis-data-cor}
# Export an empty correlation matrix
export_empty_cor_matrix(
  imported_marginals,
  folder_path = folder_path
)
# Reimport the matrix
cor_matrix <- import_cor_matrix(file.path(folder_path, "correlation_matrix.csv"))
# Add assumed correlations (note the symmetry)
cor_matrix["RATRIAL_N", "STRK14"] <- -0.02
cor_matrix["STRK14", "RATRIAL_N"] <- -0.02

# Synthesise data specifying the correlation matrix
sim_df_cor <- synthesise_data(imported_marginals, correlation_matrix = cor_matrix)
```

# Summarise the synthesised data
Summarise the synthesised data (with correlations)
```{r summarise-sim-data-cor}
# Convert to Factors from Character variables
sim_df_cor <- sim_df_cor |> dplyr::mutate_if(is.character, factor)
# Summarise the synthesised dataset
summary(sim_df_cor)
```
# Fit a cox model on the simulated data with correlations
Using the synthesised data (with correlations) fit a Cox PH model, with the same parameters as earlier.
```{r sim-data-model-cor}

# Again events are measured at 14 days
sim_df_cor$DAY <- 14

# Again check that the missing values where added
sum(sim_df_cor$RATRIAL == "")

# Again remove the missing values
sim_df_cor <- sim_df_cor[!sim_df_cor$RATRIAL == "",]

# Again drop the missing factor
sim_df_cor$RATRIAL <- droplevels(sim_df_cor$RATRIAL)

# Show there are no missing values
summary(sim_df_cor$RATRIAL)

# Fit the model on the synthesised data (with correlations)
cox.ph.sim.cor <- coxph(Surv(DAY, STRK14) ~ AGE + SEX + RATRIAL + RSBP, data = sim_df_cor)

# Show a summary of the model
cox.ph.sim.cor
```

